Community-Conditioned Language Models

### 2020-11-12 paper/paper.tex
addressing some comments

Linguistic communication requires that speakers share 
knowledge of certain linguistic conventions, such as syntactic structure, 
word meaings, and accepted patterns of interaction.
Speakers assume that these conventions are common ground based on joint membership
in some community \cite{Clark1996}.
Such a community may range in size and specificity from the very large and general
such as \emph{speakers of a language} to very small and specific, such as 
of a dialectical region

\emph{common ground} \citep{Stalnaker2002}

This basis allows for variation in linguistic convention accross communities.

The \emph{speech community} is an important concept 
in the study of linguistic variation.
Any social group exhibiting some degree of linguistic idiosyncrasy
may be considered a speech community \citep{Gumperz1972}.

Broadly construed, we want to investigate two questions.  First, we
are interested in variation in the level of linguistic complexity
exhibited by different communities.

@@

It is folklore that certain communities are more complex
linguistically: here we want to quantify this phenomenon.

More precisely, we investigate if this degree of complexity be
captured in a computational model.  For this purpose define various
Community Conditioned Language Models (CCLM for short), which can be
modeled as an estimator $M$ what a given message $m$ is acceptable in
a given community $c$: \(P_M(m \mid c)\).

@@

To investigate variation across communities, we use comments collected
from the social media website Reddit.\footnote{Comments were obtained
  from the archive at \href{https://pushshift.io/}{pushshift.io}
  \cite{Baumgartner2020}.}

@@

We experiment with two kinds of architectures for our models: an LSTM
and a Transformer architecture\jp{citations}.  In either case, they
are organised as a standard $n$-layer neural sequence encoder with
token embedding (we set $n=3$ for all models). As standard, words are predicted
using a softmax projection layer.
Such encoders form the
core of the community-conditioned language models (CCLM).

To condition the model, we add a \emph{community embedding} parameter to the models.
This parameter is concatenated with the hidden layer of the sequence encoder,
for some layer $c \leq n$, and passed through a linear layer 
which projects the resulting vector back to the origial hidden layer size.
For $c = n$, the output of this linear layer is passed directly to the softmax function,
just as the final hidden layer of the sequence encoder is in other models.
For $c=0$\jp{it's unfortunate that the letter $c$ is clashing. To Fix!}, the community embedding is concatenated with the token embedding.
For this reason, we set the hidden size of the sequence encoder 
and the size of the token to be embedding equal for all models.

@@

We combined the 46 community corpora and split the combined corpus,
reserving 10\% for testing, 10\% for validation and the rest for training.
Splits were stratified by community so that the same number of training examples
were available for each community.\jp{is it the same for testing/validation? How do you maintain this split?}

We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consists of the most frequent tokens at the message level,
meaning that we only count each token once per message.
The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
many times in a single message.

We trained the models on a simple left-to-right\jp{I don't think that this applies for the transformer model} language modeling task 
with cross entropy loss over the vocabulary 
and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

All experiments use models with \num{3} sequence encoder layers,
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen in part to give the LSTM and transformer
  models a comparable number of paramters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions.

@@

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths .
For the LSTM encoder, 
the best model ($c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

@@

We also note that the conditioned models perform differently across different communities---\
for some communities, the test \jp{and training?}perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},
whereas for others it is the same or even slightly higher.
\bn{TODO: talk about this in terms of mutual information instead?}
As an anecdotal observation, we note 
that communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics.\jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech\jp{can you say speech for written language?} patterns.

@@

\begin{table*}
  \footnotesize
  \centering
  \input{floats/model_ppls.tex}
  \caption{Mean model test perplexity per community.}
  \label{tab:model-ppls}
\end{table*}

>>>>

Linguistic communication requires that speakers share 
knowledge of certain linguistic conventions, such as syntactic structure, 
word meanings, and accepted patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among their interlocutors,
based on joint membership in a community \cite{Stalnaker2002, Clark1996}.
Such a community may be of any size,
from the very small, like members of a particular friend group,
to the very large, like speakers of English.
Because it is the source of linguistic convention, the \emph{speech community} 
is an important concept in the study of linguistic variation \citep{Gumperz1972}.

In this work, we investigate linguistic variation across online communities.
Broadly construed, we want to investigate two questions.  First, we
are interested in variation in the level of linguistic complexity
exhibited by different communities.

@@

It is folklore that certain communities are more linguistically complex than others.
Here we want to quantify this phenomenon.
More precisely, we investigate if the degree of complexity can be
captured in a computational model.  For this purpose define various
Community Conditioned Language Models (CCLM for short), which can be
modeled as an estimator $M$ of the degree to which a given message $m$ 
is acceptable in a given community $c$: \(P_M(m \mid c)\).

@@

To investigate variation across communities, we use comments collected
from the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}.}

@@

We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}
and Transformer \citep{Vaswani2017}.  In either case, the model
is organised as a standard $n$-layer neural sequence encoder with
token embedding (we set $n=3$ for all models). As is standard, words are predicted
using a softmax projection layer over a pre-determined vocabulary.
These encoders form the core of the community-conditioned language models (CCLM).

To condition the model, we add a \emph{community embedding} parameter to the models.
This parameter is concatenated with the hidden layer of the sequence encoder,
for some layer $l_c \leq n$, and passed through a linear layer 
which projects the resulting vector back to the original hidden layer size.
For $l_c = n$, the output of this linear layer is passed directly to the softmax function,
just as the final hidden layer of the sequence encoder is in other models.
For $l_c=0$, the community embedding is concatenated with the token embedding.
For this reason, we set the hidden size of the sequence encoder 
and the size of the token to be embedding equal for all models.

@@

We combined the 46 community corpora and split the combined corpus,
reserving 10\% for testing, 10\% for validation and the rest for training.
Splits were stratified by community so that each community had \num{5000} examples
for testing, \num{5000} for evaluation, and \num{40000} for training.

We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
The vocabulary consists of the most frequent tokens at the message level (across all communities),
meaning that we only count each token once per message.
The main reason for this is to avoid over-counting otherwise rare tokens that were spammed
many times in a single message.

We trained the models on a simple left-to-right language modeling task
with cross entropy loss over the vocabulary. 
Since the transformer operates on all tokens in the sequence at once,
this was achieved by masking and incrementally un-masking input tokens
in a left-to-right fashion. \bn{does this address it?} %\jp{I don't think that this applies for the transformer model} 
We used the AdamW \citep{Loshchilov2019} optimization algorithm, 
with an initial learning rate of \num{0.001}.
We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
During training, a dropout rate of $0.1$ was applied between encoder layers 
and after each linear layer.

All experiments use models with \num{3} sequence encoder layers,
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen in part to give the LSTM and transformer
  models a comparable number of paramters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions. 
We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).

@@

As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths .
For the LSTM encoder, 
the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($l_c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

@@

We also note that the conditioned models perform differently across different communities---\
for some communities, perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},\bn{is this ok took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
whereas for others it is the same or even slightly higher.
\bn{TODO: talk about this in terms of mutual information instead?}
As an anecdotal observation, we note 
that communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics,
such as /r/relationships, where the subject matter is relevant to a broad
range of people. \bn{I really mean this as a quality of the topic, rather than the community itself.}% \jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.

This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}

@@

\begin{table*}
  \footnotesize
  \centering
  \input{floats/model_ppls.tex}
  \caption{Mean model perplexity for messages in the test set, broken down by community.}
  \label{tab:model-ppls}
\end{table*}
